### 2022-01-12 src/tex/ms.tex
starting from scratch

% Define document class
\documentclass[twocolumn]{aastex631}

% Begin!
\begin{document}

% Title
\title{\showyourwork: a workflow for open source scientific articles}

% Author list
\author[0000-0002-0296-3826]{Rodrigo Luger}
\author{Others TBD}

% Abstract with filler text
\begin{abstract}
    Abstract coming soon.
\end{abstract}

% Main body with filler text
\section{Introduction}
\label{sec:intro}

\begin{figure}[ht!]
    \begin{centering}
        \includegraphics[width=\linewidth]{figures/eccentricity.pdf}
        \caption{
            The effect of binary eccentricity on the detectability of a \emph{LISA} gravitational wave source; reproduced from Figure 3 in \citet{Wagg2021}. 
            This figure was automatically generated from the script \texttt{src/figures/eccentricity.py}.
        }
        \label{fig:eccentricity}
    \end{centering}
\end{figure}

\begin{figure}[th!]
    \begin{centering}
        \includegraphics[width=\linewidth]{figures/banner.png}
        \caption{
            The default \texttt{README.md} banner in a repository instantiated from the \texttt{showyourwork-template}, with links to the \texttt{GitHub Action} build logs, a tarball containing the \texttt{TeX} source for the article, a directed acyclic graph (DAG) of the build process, and the compiled article PDF, respectively.
            Note that since this figure is a screenshot, we place it in the \texttt{src/static} folder and \texttt{git}-commit it.
            The workflow skips the generation step for any figures in that directory; note the absence of a margin icon next to this caption.
        }
        \label{fig:banner}
    \end{centering}
\end{figure}

\bibliography{bib}

\end{document}