### 2022-01-14 src/tex/ms.tex
beauty

% Define document class
\documentclass[twocolumn]{aastex631}
\usepackage{showyourwork}

% Begin!
\begin{document}

% Title
\title{A workflow for open source scientific articles}

% Author list
\author[0000-0002-0296-3826]{Rodrigo Luger}
\author{Others TBD}

% Abstract with filler text
\begin{abstract}
    Abstract coming soon.
\end{abstract}

% Main body with filler text
\section{Introduction}
\label{sec:intro}

\begin{figure}[ht!]
    \begin{centering}
        \includegraphics[width=\linewidth]{figures/eccentricity.pdf}
        \caption{Foo}
        \label{fig:eccentricity}
    \end{centering}
\end{figure}

\begin{figure}[th!]
    \begin{centering}
        \includegraphics[width=\linewidth]{figures/banner.png}
        \caption{FOO}
        \label{fig:banner}
    \end{centering}
\end{figure}

\bibliography{bib}

\end{document}